### 2021-03-18 text/text.tex
add abstract
fix typos

Each Edoc item has 43 main data fields. Most of these data fields belong
to the class of descriptive metadata. The data entry for an item's
descriptive metadata is undertaken by the person who uploads the item
which is (in most cases) one of the item's authors. Note that an item is
first uploaded to the University of Basel's research database
\url{https://www.forschdb2.unibas.ch} before being automatically fed
into Edoc. There are explicit rules on how to enter the data as
summarized in the user manual (Universität Basel, n.d.). However, there
is no (or at least no systematic) manual or automatic validation of the
entered data. This means that some data fields are very heterogenous and
require further parsing (see \protect\hyperlink{gold-standard}{section
``Gold standard''}). The data fields that are of interest to the project
at hand are discussed in more detail in (see
\protect\hyperlink{sample-data-set}{section ``Sample data set''} below).

@@

In this section, I will explain how the sample data set is selected and
constructed. ``Selection'' hereby means the task of specifying a subset
of the Edoc data; by ``construction'' I mean the task of implementing
this selection.

@@

Taking into account the third constraint simply means that we do not
employ any further restrictions. We can hence construct the Edoc sample
data set by choosing exactly those items from
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}}
which have non-empty data fields \texttt{title}, \texttt{abstract},
\texttt{id\_number}, and \texttt{keywords}. We do so by calling
\texttt{Data.select\_from\_file} iteratively for all files in
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}}.
The resulting file is saved in
\href{https://github.com/MHindermann/mas/tree/main/files/sample}{\texttt{/sample/}}
as \texttt{sample\_master.json}.

@@

Of the 68'345 items in
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}},
all have a title (non-empty \texttt{title} data field), a little more
than half of the items have an abstract (37'381 items with non-empty
\texttt{abstract} data field), roughly half of the items have an ID
(35'355 items with non-empty \texttt{id\_number} data field),\footnote{Note
  that when using the facet by blank on \texttt{id\_number} in
  OpenRefine, there are 57'153 matches; this is due to the fact that
  many items with an ID such as DOI have secondary or tertiary IDs such
  as ISI or PMID.} and less than 10\% of the items have keywords (6'660
items with non-empty \texttt{keywords} data field). The Edoc sample data
set as requires all the above data fields to be non-empty;
\texttt{/sample/sample\_master.json} has 4'111 items and hence
constitutes 6\% of the raw data.

@@

Annif (Suominen et al. 2021) is an open source multi-algorithm automated
indexing tool developed at the Finnish National Library (see Suominen
2019 for an overview). Training a local Annif instance for machine
indexing is currently evaluated at different libraries, for example at
the German National Library (see Nagelschmidt 12.11.2020). As stated in
section \protect\hyperlink{introduction}{section ``Introduction''}, in
this project I am interested in using Annif out of the box. This means
that instead of training a local Annif instance, I will be relying on
the Annif REST-style API (see \url{https://api.annif.org/}).\footnote{This
  API is specifically intended for testing. For integration into a
  production system, the Finto AI API is available (see
  \url{https://ai.finto.fi/v1/ui/}).} The advantage of this approach is
that it is quick and relatively easy to implement. The disadvantage is
that no custom algorithms or vocabularies can be used. The task of
constructing a gold standard (see
\protect\hyperlink{gold-standard}{section ``Gold standard''}) remains
the same.

The vocabularies provided by the Annif API are the General Finnish
Ontology or YSO in short (see \url{https://finto.fi/yso/en/}) and
Wikidata (see \url{https://www.wikidata.org/}). Each vocabulary is
combined with an algorithm or algorithm ensemble respectively and a
specific language to create a so-called Annif ``project.'' The five
projects used in what follows are summarized in Table 1. A high-level
description of the algorithm(s) is given by Suominen (2019, 7--10).
Information on the implementation of the algorithm(s) can be found at
the Annif GitHub Wiki (see
\url{https://github.com/NatLibFi/Annif/wiki}).

@@

As explained in \protect\hyperlink{machine-indexing}{section ``Machine
indexing''}, Annif assigns index terms from a controlled vocabulary. If
we want to assess the quality of the indexing via a gold standard, we
must therefore ensure that the gold standard makes use of the vocabulary
used by Annif. The relevant (English) vocabularies are Wikidata and YSO.
The next step in constructing the derivative gold standard is hence to
match the extracted and cleaned keywords with keywords from Wikidata and
YSO. This process is called ``reconciliation'' (see
\url{https://docs.openrefine.org/manual/reconciling}) and the tool of
choice for this task is OpenRefine).\footnote{For data refinement I use
  OpenRefine version 3.4.1 (see \url{https://openrefine.org/}). Data
  manipulation in OpenRefine is tracked: the manipulation history of
  some data can be exported as JSON file and then reproduced (on the
  same data on a different machine or on different data) by loading said
  file. I will supply a corresponding manipulation history whenever
  appropriate. Note that when using OpenRefine with larger files (and
  there are many such files in this project), the memory allocation to
  OpenRefine must be manually increased (see
  \url{https://docs.openrefine.org/manual/installing/\#increasing-memory-allocation}
  for details). Also note that OpenRefine always counts the number of
  records by the first column (a fact which caused me many headaches).}

@@

Second, systematic biases were identified and removed. The most
prevalent bias was an due to a suggestion's type. As stated above, the
reconciliation service API was not constrained by type but had access to
the complete Wikidata database. Since Wikidata is an ontology that
encompasses everything, it also features types whose concepts cannot
qualify as subject terms (at least in the present context). The most
prominent example is the type Q13442814 ``scholarly article.'' Wikidata
contains the metadata of many scholarly articles. Now, for some of our
keywords, there is a scholarly article with a title that exactly matches
the keyword; and since there is no restriction concerning the type, the
scholarly article is suggested with high confidence (see
\url{https://github.com/OpenRefine/OpenRefine/wiki/Reconciliation-Service-API}
for details). To generalize, suggestions with types whose concepts are
proper names are usually incorrect. Based on this observation,
suggestions with types such as ``scholarly article,'' ``clinical
trial,'' and ``scientific journal'' were rejected.\footnote{The complete
  list includes ``academic journal,'' ``open access journal,''
  ``thesis,'' ``doctoral thesis,'' and ``natural number.''} Suggestions
types such as ``human,'' ``album,'' and ``commune of France'' were
manually verified (i.e., checked for correctness).\footnote{The complete
  list includes ``film,'' ``musical group,'' ``business,'' ``literary
  work,'' ``television series,'' ``organization,'' ``family name,''
  ``written work,'' ``video game,'' ``single, television series
  episode,'' ``painting,'' , ``city of the United States,''
  ``magazine,'' ``studio album,'' ``year,'' ``nonprofit organization,''
  ``border town,'' ``international organization,'' ``political party,''
  ``software,'' ``song,'' ``website,'' ``article comic strip,''
  ``collection,'' ``commune of Italy,'' ``fictional human,'' ``film,''
  ``government agency,'' ``village,'' ``academic journal article,''
  ``female given name,'' and ``poem.''}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  In response to the result that increasing the text basis to titles and
  abstracts does not increase performance discussed in
  \protect\hyperlink{assessment}{section ``Assessment''}, add full text
  support and evaluate the performance by increasing the text basis even
  further. To do this, \texttt{Data.enrich\_with\_annif} must be amended
  to fetch and parse the full text PDFs. For an item in the sample data
  set, the link to the full text is constructed from the data fields
  \texttt{offical\ url} and \texttt{documents\ -\ main}. Initial tests
  using the Tika Python library for parsing were promising.
\item
  Originally I had intended to undertake an additional performance test
  based on a gold standard distinct from the derivative gold standard
  constructed in \protect\hyperlink{gold-standard}{section ``Gold
  standard''}. In order to do so, items in the Edoc sample data set were
  enriched with MeSH terms (see
  \url{https://www.nlm.nih.gov/mesh/meshhome.html}) retrieved from
  PubMed (if they had a PubMed ID). This was achieved by calling
  \texttt{Data.enrich\_with\_mesh} on
  \texttt{/indexed/indexed\_master.json}. However, the Annif API does
  not include a configuration with a MeSH vocabulary, and there is no
  concordance between MeSH and either Wikidata or YSO (see
  \url{https://coli-conc.gbv.de/cocoda/app/concordances.html}).
  Therefore, the MeSH terms obtained from PubMed would have to be
  reconciled with Wikidata and YSO in a very time-consuming manner,
  yielding again only a derivative gold standard. As a way forward, it
  would be more effective to train a local instance of Annif on the
  already obtained MeSh data.
\item
  As shown in \protect\hyperlink{assessment}{section ``Assessment''},
  the performance of Annif configurations varies with department. It is
  hence plausible that some (subsets of) items in Edoc require
  domain-specific vocabularies that are not available in Annif. These
  vocabularies can be identified using BARTOC FAST (Hindermann and Ledl
  2020); the implementation is straightforward with my bartocsuggest
  Python library that already includes an Annif wrapper (see
  \url{https://pypi.org/project/bartocsuggest/}).
\item
  An issue that has been neglected in this project is the fact that Edoc
  is a multilingual repository containing more than 3'324 items in
  German (roughly 5\% of Edoc) and 273 items in other languages (less
  than 0.5\% of Edoc).\footnote{The Edoc sample data set only contains
    53 non-English (inclduding 48 German) items; the negligence is hence
    well justified.} Here it would be interesting to train a local
  instance of Annif with GND similar in spirit to Nagelschmidt
  (12.11.2020) and compare its perfomance to other machine indexing
  tools such as DA-3 (Beckmann et al. 2019) which are currently
  informally evaluated at the University Library Basel.
\end{enumerate}

Finally, I had initially {[}monitoring,,,{]}.

As mentioned in \protect\hyperlink{introduction}{section
``Introduction''}, all files, data, and code used in this project is
available at \url{https://github.com/MHindermann/mas} and will not be
replicated here.

\hypertarget{bibliography}{%
\section{Bibliography}\label{bibliography}}

@@

>>>>

Each Edoc item has 43 main data fields. Most of these data fields belong
to the class of descriptive metadata. The data entry for an item's
descriptive metadata is undertaken by the person who uploads the item
which is (in most cases) one of the item's authors. Note that an item is
first uploaded to the University of Basel's research database
(\url{https://www.forschdb2.unibas.ch}) before being automatically fed
into Edoc. There are explicit rules on how to enter the data as
summarized in the user manual (Universität Basel, n.d.). However, there
is no (or at least no systematic) manual or automatic validation of the
entered data. This means that some data fields are very heterogenous and
require further parsing (see \protect\hyperlink{gold-standard}{section
``Gold standard''}). The data fields that are of interest to the project
at hand are discussed in more detail in
\protect\hyperlink{sample-data-set}{section ``Sample data set''} below.

@@

In this section, I will explain how the sample data set was selected and
constructed. ``Selection'' hereby means the task of specifying a subset
of the Edoc data; by ``construction'' I mean the task of implementing
this selection.

@@

Taking into account the third constraint simply means that we do not
employ any further restrictions. We can hence construct the Edoc sample
data set by choosing exactly those items from
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}}
which have non-empty data fields \texttt{title}, \texttt{abstract},
\texttt{id\_number}, and \texttt{keywords}. We do so by calling
\texttt{Data.select\_from\_file} iteratively for all files in
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}}
using the aforementioned data fields as parameters. The resulting file
is saved in
\href{https://github.com/MHindermann/mas/tree/main/files/sample}{\texttt{/sample/}}
as \texttt{sample\_master.json}.

@@

Of the 68'345 items in
\href{https://github.com/MHindermann/mas/tree/main/files/raw}{\texttt{/raw/}},
all have a title (non-empty \texttt{title} data field), a little more
than half of the items have an abstract (37'381 items with non-empty
\texttt{abstract} data field), roughly half of the items have an ID
(35'355 items with non-empty \texttt{id\_number} data field),\footnote{Note
  that when using the facet by blank on \texttt{id\_number} in
  OpenRefine, there are 57'153 matches; this is due to the fact that
  many items with an ID such as DOI have secondary or tertiary IDs such
  as ISI or PMID.} and less than 10\% of the items have keywords (6'660
items with non-empty \texttt{keywords} data field). The Edoc sample data
set requires all the above data fields to be non-empty;
\texttt{/sample/sample\_master.json} has 4'111 items and hence
constitutes 6\% of the raw data.

@@

Annif (Suominen et al. 2021) is an open source multi-algorithm automated
indexing tool developed at the Finnish National Library (see Suominen
2019 for an overview). Training a local Annif instance for machine
indexing is currently evaluated at different libraries, for example at
the German National Library (see Nagelschmidt 12.11.2020). As stated in
\protect\hyperlink{introduction}{section ``Introduction''}, in this
project I am interested in using Annif out of the box. This means that
instead of training a local Annif instance, I will be relying on the
Annif REST-style API (see \url{https://api.annif.org/}).\footnote{This
  API is specifically intended for testing. For integration into a
  production system, the Finto AI API is available (see
  \url{https://ai.finto.fi/v1/ui/}).} The advantage of this approach is
that it is quick and relatively easy to implement. The disadvantage is
that no custom algorithms or vocabularies can be used. The task of
constructing a gold standard (see
\protect\hyperlink{gold-standard}{section ``Gold standard''}) remains
the same.

The vocabularies provided by the Annif API are the General Finnish
Ontology or YSO in short (see \url{https://finto.fi/yso/en/}) and
Wikidata (see \url{https://www.wikidata.org/}). Each vocabulary is
combined with an algorithm or algorithm ensemble respectively and a
specific language to create a so-called Annif ``project.'' The five
projects used in what follows are summarized in Table 1. A high-level
description of the algorithm(s) is given by Suominen (2019, 7--10);
information on the implementation of the algorithm(s) can be found at
the Annif GitHub Wiki (see
\url{https://github.com/NatLibFi/Annif/wiki}).

@@

As explained in \protect\hyperlink{machine-indexing}{section ``Machine
indexing''}, Annif assigns subject terms from a controlled vocabulary.
If we want to assess the quality of the indexing via a gold standard, we
must therefore ensure that the gold standard makes use of the vocabulary
used by Annif. The relevant (English) vocabularies are Wikidata and YSO.
The next step in constructing the derivative gold standard is hence to
match the extracted and cleaned keywords with keywords from Wikidata and
YSO. This process is called ``reconciliation'' (see
\url{https://docs.openrefine.org/manual/reconciling}) and the tool of
choice for this task is OpenRefine).\footnote{For data refinement I use
  OpenRefine version 3.4.1 (see \url{https://openrefine.org/}). Data
  manipulation in OpenRefine is tracked: the manipulation history of
  some data can be exported as JSON file and then reproduced (on the
  same data on a different machine or on different data) by loading said
  file. I will supply a corresponding manipulation history whenever
  appropriate. Note that when using OpenRefine with larger files (and
  there are many such files in this project), the memory allocation to
  OpenRefine must be manually increased (see
  \url{https://docs.openrefine.org/manual/installing/\#increasing-memory-allocation}
  for details). Also note that OpenRefine always counts the number of
  records by the first column (a fact which caused me many headaches).}

@@

Second, systematic biases were identified and removed. The most
prevalent bias was due to a suggestion's type. As stated above, the
reconciliation service API was not constrained by type but had access to
the complete Wikidata database. Since Wikidata is an ontology that
encompasses everything, it also features types whose concepts cannot
qualify as subject terms (at least in the present context). The most
prominent example is the type Q13442814 ``scholarly article.'' Wikidata
contains the metadata of many scholarly articles. Now, for some of our
keywords, there is a scholarly article with a title that exactly matches
the keyword; and since there is no restriction concerning the type, the
scholarly article is suggested with high confidence (see
\url{https://github.com/OpenRefine/OpenRefine/wiki/Reconciliation-Service-API}
for details). To generalize, suggestions with types whose concepts are
proper names are usually incorrect. Based on this observation,
suggestions with types such as ``scholarly article,'' ``clinical
trial,'' and ``scientific journal'' were rejected.\footnote{The complete
  list includes ``academic journal,'' ``open access journal,''
  ``thesis,'' ``doctoral thesis,'' and ``natural number.''} Suggestions
types such as ``human,'' ``album,'' and ``commune of France'' were
manually verified (i.e., checked for correctness).\footnote{The complete
  list includes ``film,'' ``musical group,'' ``business,'' ``literary
  work,'' ``television series,'' ``organization,'' ``family name,''
  ``written work,'' ``video game,'' ``single, television series
  episode,'' ``painting,'' , ``city of the United States,''
  ``magazine,'' ``studio album,'' ``year,'' ``nonprofit organization,''
  ``border town,'' ``international organization,'' ``political party,''
  ``software,'' ``song,'' ``website,'' ``article comic strip,''
  ``collection,'' ``commune of Italy,'' ``fictional human,'' ``film,''
  ``government agency,'' ``village,'' ``academic journal article,''
  ``female given name,'' and ``poem.''}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  In response to the result that increasing the text basis to titles and
  abstracts does not increase performance discussed in
  \protect\hyperlink{assessment}{section ``Assessment''}, add full text
  support and evaluate the performance by increasing the text basis even
  further. To do this, \texttt{Data.enrich\_with\_annif} must be amended
  to fetch and parse the full text PDFs. For an item in the sample data
  set, the link to the full text is constructed from the data fields
  \texttt{offical\ url} and \texttt{documents\ -\ main}. Initial tests
  using the Tika Python library for parsing were promising.
\item
  Originally I had intended to undertake an additional performance test
  based on a gold standard distinct from the derivative gold standard
  constructed in \protect\hyperlink{gold-standard}{section ``Gold
  standard''}. In order to do so, items in the Edoc sample data set were
  enriched with MeSH terms (see
  \url{https://www.nlm.nih.gov/mesh/meshhome.html}) retrieved from
  PubMed (if they had a PubMed ID). This was achieved by calling
  \texttt{Data.enrich\_with\_mesh} on
  \texttt{/indexed/indexed\_master.json}. However, the Annif API does
  not include a configuration with a MeSH vocabulary, and there is no
  concordance between MeSH and either Wikidata or YSO (see
  \url{https://coli-conc.gbv.de/cocoda/app/concordances.html}).
  Therefore, the MeSH terms obtained from PubMed would have to be
  reconciled with Wikidata and YSO in a very time-consuming manner,
  yielding again only a derivative gold standard. As a way forward, it
  would be more effective to train a local instance of Annif on the
  already obtained MeSH data.
\item
  As shown in \protect\hyperlink{assessment}{section ``Assessment''},
  the performance of Annif configurations varies with department. It is
  hence plausible that some (subsets of) items in Edoc require
  domain-specific vocabularies that are not available in Annif. These
  vocabularies can be identified using BARTOC FAST (Hindermann and Ledl
  2020); the implementation is straightforward with my bartocsuggest
  Python library that already includes an Annif wrapper (see
  \url{https://pypi.org/project/bartocsuggest/}).
\item
  An issue that has been neglected in this project is the fact that Edoc
  is a multilingual repository containing more than 3'324 items in
  German (roughly 5\% of Edoc) and 273 items in other languages (less
  than 0.5\% of Edoc).\footnote{The Edoc sample data set only contains
    53 non-English (inclduding 48 German) items; the negligence is hence
    well justified.} Here it would be interesting to train a local
  instance of Annif with GND similar in spirit to Nagelschmidt
  (12.11.2020) and compare its perfomance to other machine indexing
  tools such as DA-3 (Beckmann et al. 2019) which are currently
  informally evaluated at the University Library Basel.
\end{enumerate}

\hypertarget{bibliography}{%
\section*{Bibliography}\label{bibliography}}
\addcontentsline{toc}{section}{Bibliography}

@@

See \url{https://github.com/MHindermann/mas}.
